---
title: "Assignment 1"
author: "Ella Lee"
format: html
editor: visual
---

## Preprocessing

```{r}
install.packages("wordcloud")
install.packages("SnowballC")

library(rtweet)
library(tm)
library(wordcloud)
library(SnowballC)

```

Set 1

```{r}
set1_data <- read.csv("bshes740/TADA_A1_set1.csv")
set1_texts <- set1_data$text
summary(set1_texts)

set1_text_corpus <- Corpus(VectorSource(set1_texts))
set1_text_corpus <- tm_map(set1_text_corpus, content_transformer(tolower))
set1_text_corpus <- tm_map(set1_text_corpus, removeNumbers)
stopwords()


set1_text_corpus <- tm_map(set1_text_corpus, stemDocument)

set1_text_corpus[[1]]
set1_text_corpus[[1]]$content
set1_text_corpus[[1]]$meta$datetimestamp
doc_term_matrix <- DocumentTermMatrix(set1_text_corpus)
doc_term_matrix

```

Set 2

```{r}
set2_data <- read.csv("bshes740/TADA_A1_set2.csv")

set2_texts <- set2_data$text
summary(set2_texts)

set2_text_corpus <- Corpus(VectorSource(set2_texts))
set2_text_corpus <- tm_map(set2_text_corpus, content_transformer(tolower))
set2_text_corpus <- tm_map(set2_text_corpus, removeNumbers)
stopwords()

set2_text_corpus <- tm_map(set2_text_corpus, stemDocument)

set2_text_corpus[[1]]
set2_text_corpus[[1]]$content
set2_text_corpus[[1]]$meta$datetimestamp
doc_term_matrix2 <- DocumentTermMatrix(set2_text_corpus)
doc_term_matrix2


```

## 1. date range, language

What are the date ranges for the two sets? What information is provided in the CSV files? What are the languages in which tweets have been posted?

```{r}
#Date range

date_range_set1 <- range(set1_data$date, na.rm = TRUE)
print(paste("Set 1 Date Range:", date_range_set1[1], "to", date_range_set1[2]))
##"Set 1 Date Range: 2019-11-01 to 2020-01-30"

date_range_set2 <- range(set2_data$date, na.rm = TRUE)
print(paste("Set 2 Date Range:", date_range_set2[1], "to", date_range_set2[2]))
##"Set 2 Date Range: 2020-02-01 to 2020-04-29"

#information in the CSV file
colnames(set1_data)
summary(set1_data)
colnames(set2_data)
summary(set2_data)
##The CSV files provide tweet IDs, dates, languages, tweet content, and location

#language usage
language_counts <- table(set1_data$lang)
print(language_counts)
language_counts2 <- table(set2_data$lang)
print(language_counts2)

##Set 1: ar, ca, cs, cy, da, de, el, en, es, et, eu, fi, fr, hi, ht, hu, in, is, it, iw, ja, ko, lt, lv, mr, nl, no, pl, pt, ro, sl, sv, th, tl, tr, und, vi, zh

##Set 2: ar, ca, cs, cy, da, de, el, en, es, et, eu, fa, fi, fr, hi, ht, hu, in, is, it, ja, ko, lt, lv, nl, no, pl, pt, ro, ru, sl, sv, th, tl, tr, und, vi, zh

```

## 2. total number

What is the total number of posts in set 1? What is the total in set 2?

```{r}
total_posts_set1 <- nrow(set1_data)
print(paste("Total number of posts in Set 1:", total_posts_set1))
total_posts_set2 <- nrow(set2_data)
print(paste("Total number of posts in Set 2:", total_posts_set2))

##"Total number of posts in Set 1: 171012"
##"Total number of posts in Set 2: 226852"
```

## 3. number of tweets

How many tweets are there for methadone, Suboxone, and fentanyl in total? Tip: sometimes alternative expressions are used for substances (eg., fent for fentanyl).

```{r}
# Extract the 'text' column from both CSV datasets as vectors
set1_texts <- set1_data$text
set2_texts <- set2_data$text

# Methadone
set1_methadone <- grep("methadone", tolower(set1_texts))
set2_methadone <- grep("methadone", tolower(set2_texts))

# Suboxone
set1_suboxone <- grep("suboxone", tolower(set1_texts))
set2_suboxone <- grep("suboxone", tolower(set2_texts))


# Fentanyl and "fent"
set1_fent <- grep("fentanyl|fent", tolower(set1_texts))
set2_fent <- grep("fentanyl|fent", tolower(set2_texts))


# Calculate totals
total_methadone <- length(set1_methadone) + length(set2_methadone)
total_suboxone <- length(set1_suboxone) + length(set2_suboxone)
total_fentanyl <- length(set1_fent) + length(set2_fent)

#print
total_methadone
##21219
total_suboxone
##9324
total_fentanyl
##353417


```

## 4. fentanyl analogs

Are there fentanyl analogs that are also being discussed (eg., carfentanil)?

```{r}
# Define a vector of fentanyl analogs to search for
fentanyl_analogs <- c("carfentanil")

# Initialize lists to store results
set1_analog <- list()
set2_analog <- list()

# Loop through the analogs and grep for each one
for (analog in fentanyl_analogs) {
  set1_analog[[analog]] <- grep(analog, tolower(set1_texts))
  set2_analog[[analog]] <- grep(analog, tolower(set2_texts))
}

# Print the results
for (analog in fentanyl_analogs) {
  total_set1 <- length(set1_analog[[analog]])
  total_set2 <- length(set2_analog[[analog]])
}  
 total_set1
 ## 854
 total_set2
 ## 430
 
```

## 5. most associated top 5-10

What are some of the topics that are most closely associated with each of the three substances? The top 5-10 topics (if relevant) are acceptable.

```{r}
#methadone
methadone_assoc <- findAssocs(doc_term_matrix, "methadone", 0.3)
methadone_assoc2 <- findAssocs(doc_term_matrix2, "methadone", 0.3) 

#suboxone
suboxone_assoc <- findAssocs(doc_term_matrix, "suboxone", 0.3)
suboxone_assoc2 <- findAssocs(doc_term_matrix2, "suboxone", 0.3) 

#fentanyl
fentanyl_assoc <- findAssocs(doc_term_matrix, "fentanyl", 0.3)
fentanyl_assoc2 <- findAssocs(doc_term_matrix2, "fentanyl", 0.3) 


```

## 6. word clouds

Generate word clouds for each set, so that they can be shown to the researcher.

```{r}
set1_matrix <- as.matrix(doc_term_matrix)
set1_words <- sort(rowSums(set1_matrix), decreasing = TRUE)
set1_word_data <- data.frame(word = names(set1_words), freq = set1_words)

set2_matrix <- as.matrix(doc_term_matrix2)
set2_words <- sort(rowSums(set2_matrix), decreasing = TRUE)
set2_word_data <- data.frame(word = names(set2_words), freq = set2_words)

set1_wordcloud <- wordcloud(set1_word_data$word,
                            set1_word_data$freq, 
                            min.freq = 3, 
                            max.words = 100, 
                            random.order = FALSE, 
                            rot.per = 0.35, 
                            scale = c(3, 0.5))


set2_wordcloud <- wordcloud(set2_word_data$word,
                            set2_word_data$freq, 
                            min.freq = 3, 
                            max.words = 100, 
                            random.order = FALSE, 
                            rot.per = 0.35, 
                            scale = c(3, 0.5))

#or
terms_with_high_freq <- findFreqTerms(doc_term_matrix, lowfreq = 15, highfreq = Inf)
summary(terms_with_high_freq)
terms_with_high_freq[1:10]
wordcloud(tweet_text_corpus, min.freq = 10, max.words = 200, scale=c(3,.1), random.order = FALSE, colors = brewer.pal(12, "Set3"))

```

## 7. time series figures

Generate appropriate time-series ï¬gures to compare how the frequencies of mentions of these substances differ.

```{r}

```

## 8. top 10 most frequent bigrams

Find the top 10 most frequent bigrams in each of the three sets. Plot a bar chart for these.
